---
title: "ecopmodb"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This package provides functionality for navigating the prediction files produced by the [ecopmo](https://github.com/BigelowLab/ecopmo) package workflows.

# 1 Requirements

## From CRAN

  + [rlang](https://CRAN.R-project.org/package=rlang)
  + [stars](https://CRAN.R-project.org/package=stars)
  + [readr](https://CRAN.R-project.org/package=readr)
  + [dplyr](https://CRAN.R-project.org/package=dplyr)
  + [stringr](https://CRAN.R-project.org/package=stringr)
  + [yaml](https://CRAN.R-project.org/package=yaml)

# 2 Installation

Use the [pak R package](https://CRAN.R-project.org/package=pak) to install directly from Github.

```
pak::pak("BigelowLab/ecopmodb")
```

# 3 Summary

Prediction output are stored in a one-prediction-per-file architecture. Depending upon the data set, there may be many files per day (or month). To facilitate the navigation, selection and reading of files we provide a simple database which can be filtered to select files, and a function that will read the files into a [stars](https://CRAN.R-project.org/package=stars) or [SpatRaster](https://CRAN.R-project.org/package=terra) array.

# 4 Read in a database

To read a database, you need only need to provide a path to where the database is located.

```{r}
suppressPackageStartupMessages({
  library(ecopmodb)
  library(dplyr)
  library(stars)
})

# obviously, you'll want to use a path that is available to you
path = "/mnt/s1/projects/ecocast/projectdata/ecomoncopernicusmodel/copernicus/ctenop/versions/v0/v0.00/preds"
db = ecopmodb::read_database(path)
dplyr::count(db, per, type)
```


## Have one place for all ecopmo outputs?

If you have one directory where all of your [ecopmo](https://github.com/BigelowLab/ecopmo) prediction outputs are stored, then you can set up a hidden file to store that path. This is just a convenience for the user, and is not a requirement.

```{r}
ecopmo::set_root_data_path("/mnt/s1/projects/ecocast/projectdata/ecomoncopernicusmodel/copernicus")
```

This will be useful if you use configuration files (see below).

## Select a subset of files

Filtering is quite easy using [dplyr](https://CRAN.R-project.org/package=dplyr).

```{r}
db = dplyr::filter(db,
                   per == "monc",
                   type == "q050")
```

## Finally, read in the files.

```{r}
x = ecopmodb::read_ecopmo(db, path)
x
```
These can be read in as [SpatRaster](https://CRAN.R-project.org/package=terra) class objects, but the arrangement of the layers may not be what you might hope.

```{r}
y = ecopmodb::read_ecopmo(db, path, form = "SpatRaster")
y
```

# 5 Using configuration lists

Configuration lists, stored in [YAML](https://yaml.org/) format, play a central role in the [ecopmo](https://github.com/BigelowLab/ecopmo) workflow.

Assuming you set the root data path (see above) it is very easy to read in a configuration.

```{r config}
cfg = ecopmodb::read_configuration(species = "ctenop", version = "v0.00")
str(cfg)
```

## Read a database with the configuration

You maybe find it easier to read a database without having to muck around with paths.
```{r read_dataase_config}
db = ecopmodb::read_database(cfg)
```